---
title: "Example_Code"
author: "Patrick D. lorch"
date: "2023-01-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Using MODISTools


```{r installload}
# install.packages("MODISTools")

library(dplyr)
library(MODISTools)
library(raster)
library(sf)
library(leaflet)
library(prism)

library(lubridate)

source("mt_to_terra.R")

dir.create("Prism")
prism_set_dl_dir("Prism")

```

## Examples from MODISTools vignette

https://cran.r-project.org/web/packages/MODISTools/vignettes/modistools-vignette.html

Values in the land cover download do not match any legend I can find.
(e.g., here: https://lpdaac.usgs.gov/documents/101/MCD12_User_Guide_V6.pdf or https://developers.google.com/earth-engine/datasets/catalog/MODIS_006_MCD12Q1#bands)

It turns out that this software is somewhat unsettled.

* steps that depend on rgdal, raster, sp, etc. will be depricated soon
* Version 1.1.3 breaks the mt_to_raster function
  * waiting on feedback from author to solve
  * solved in draft, testing

Not sure how to do legends.

```{r modisdata}
# NDVI
products <- mt_products()
View(products)
products[grep("NDVI", products$description),]
bands <- mt_bands(product = "MOD13Q1")
View(bands)
dates <- mt_dates(product = "MOD13Q1", lat = 32, lon = -114)
View(dates)
dates_june2august_ndvi = dates %>%
  filter(month(calendar_date) %in% 6:8)

# Land cover
products[grep("LC", products$description),]
bands_LC <- mt_bands(product = "MCD12Q1")
View(bands_LC)
dates_LC <- mt_dates(product = "MCD12Q1", lat = 32, lon = -114)
View(dates_LC)
dates_june2august_ndvi = dates_LC %>%
  filter(month(calendar_date) %in% 6:8)


# Set some example points
testpoints = data.frame(
  Point_id = c("Old River Channel SFWA", "Willows E of KRP", "Restoration N of KRP HQ"),
  Latitude = c(35.667130, 35.669082, 35.669857),
  Longitude = c(-118.343102, -118.301213, -118.305338))

# test code for one point
tc = mt_subset(lat= 35.667130,
               lon= -118.343102,
                product = "MOD13Q1",
                band = "250m_16_days_NDVI",
                km_lr = 0.7,
                km_ab = 0.7,
                start = "2008-06-09",
                end = "2008-06-09",
                internal = TRUE)

testpoints_sf = st_as_sf(testpoints, coords = c("Longitude", "Latitude"), crs = 4326)
View(testpoints_sf)
plot(st_geometry(testpoints_sf))

# If you need to transform
# tp_6340 = st_transform(testpoints_sf, crs = 6340)

tp_df = sfheaders::sf_to_df(testpoints_sf)
names(tp_df)[3:4] = c("lon", "lat")


# Find date of overflight for use in subsets below
dates08 = mt_dates(product = "MOD13Q1", lat = tp_df$lat[1], lon = tp_df$lon[1])
dates08[lubridate::year(dates08$calendar_date)==2008,]

subsets <- mt_batch_subset(df = tp_df,
                     product = "MOD13Q1",
                     band = "250m_16_days_NDVI",
                     # km_lr = 0.7,
                     # km_ab = 0.7,
                     start = "2008-06-09",
                     end = "2008-06-09",
                     internal = TRUE)


tp_r = mt_to_terra(subsets, reproject = T)

plot(tp_r)
plot(st_geometry(testpoints_sf), add = T)

# Find date of overflight for use in subsets below
dates08_lc = mt_dates(product = "MCD12Q1", lat = tp_df$lat[1], lon = tp_df$lon[1])
dates08_lc[lubridate::year(dates08_lc$calendar_date)==2008,]

subsets_lc <- mt_batch_subset(df = tp_df,
                     product = "MCD12Q1",
                     band = "LC_Type1",
                     # km_lr = 0.7,
                     # km_ab = 0.7,
                     start = "2008-01-01",
                     end = "2008-01-01",
                     internal = TRUE)

tp_r_lc = mt_to_terra(subsets_lc, reproject = T)

subsets_lc <- mt_batch_subset(df = tp_df,
                     product = "MCD12Q1",
                     band = "LC_Prop1",
                     km_lr = 0.7,
                     km_ab = 0.7,
                     start = "2008-01-01",
                     end = "2008-01-01",
                     internal = TRUE)

tp_r_lc = mt_to_terra(subsets_lc, reproject = T)

plot(tp_r_lc)
text(tp_r_lc)
# tp_r_lc@legend
plot(st_geometry(testpoints_sf), add = T)

```

## PRISM weather data

https://github.com/ropensci/prism

```{r prism}
prism_get_dl_dir()
# prism_set_dl_dir("Prism")
prism_check_dl_dir()

# This will take a long time, especially at SSRS
get_prism_monthlys(type = "ppt", year = 2000:2022, mon = 1:12, keepZip = FALSE)
```


## Some better plots

### Leaflet

From here:
https://rspatialdata.github.io/vegetation.html

```{r leaflet}
label_ndvi = "NDVI"
tp_r_ndvi = raster(mt_to_terra(subsets, reproject = T))

label_lc = "Land Cover Class"
tp_r_lc = raster(mt_to_terra(subsets_lc, reproject = T))

# Defining color palette
pal_ndvi <- colorNumeric(c("#440154FF", "#238A8DFF", "#FDE725FF"), values(tp_r_ndvi), na.color = "transparent")
pal_lc <- colorNumeric(c("#440154FF", "#238A8DFF", "#FDE725FF"), values(tp_r_lc), na.color = "transparent")

# Visualising using leaflet
leaflet() %>%
  addTiles() %>%
  addRasterImage(tp_r_ndvi, colors = pal_ndvi, opacity = 0.75, group = label_ndvi) %>%
  # addRasterImage(tp_r_lc, colors = pal_lc, opacity = 0.75, group = label_lc) %>%
  addLegend(
    pal = pal_ndvi, values = values(tp_r_ndvi),
    title = label_ndvi
  )  %>% 
  # addLegend(
  #   pal = pal_lc, values = values(tp_r_lc),
  #   title = label_lc
  # )  %>% 
  addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
  addProviderTiles("Esri.WorldImagery", group = "ESRI World Imagery") %>%
  addProviderTiles("Esri.WorldTopoMap", group = "ESRI World Topo") %>%
  addLayersControl(
    baseGroups = c("OpenStreetMap", "ESRI World Imagery", "ESRI World Topo"),
    # overlayGroups = c(label_ndvi, label_lc),
    overlayGroups = c(label_ndvi),
    options = layersControlOptions(collapsed = FALSE,),
    position = "bottomleft"
  )

   
```


