---
title: "Example_Code"
author: "Patrick D. lorch"
date: "2023-01-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MODIS R packages

There are 3 MODIS packages. All have problems due to transition away from packages maintained by , who is retiring, leading to the sunsetting of packages like rgdal, raster, sp, etc.

* MODIS depends on rgdal and has not been updated yet to deal with the transition
* MODIStools is being actively updated but is not fully compatible yet
  * For example, mt_to_raster needs to be replaced by mt_to_terra
  * that is why we load this modified function here
* MODIStsp is being updated but is not compatible either, still using rgdal and raster packages
  * this one has a GUI mode and a programatic mode
  

## Using MODISTools


```{r installload}
# install.packages("MODISTools")
# install.packages("MODIStsp")
# install.packages("shinydashboard")
# install.packages("shinyFiles")
# install.packages("shinyalert")
# install.packages("rappdirs")
# install.packages("shinyjs")
# install.packages("leafem")
# install.packages("mapedit")
# install.packages("magrittr")
# install.packages("MODIStsp")

# library(devtools)
# install_github(rspatial/luna)

library(dplyr)
library(MODISTools)
library(MODIStsp)
library(raster)
library(sf)
library(leaflet)
library(raster)
library(terra)
library(luna)

library(shiny)
library(shinydashboard)
library(shinyFiles)
library(shinyalert)
library(rappdirs)
library(shinyjs)
library(leafem)
library(mapedit)
library(magrittr)

library(prism)
library(rgdal)
# library(terra)
library(dismo)

library(lubridate)

source("mt_to_terra.R")

dir.create("Prism")
prism_set_dl_dir("Prism")

```

## Examples from MODISTools vignette

https://cran.r-project.org/web/packages/MODISTools/vignettes/modistools-vignette.html

Values in the land cover download do not match any legend I can find.
(e.g., here: https://lpdaac.usgs.gov/documents/101/MCD12_User_Guide_V6.pdf or https://developers.google.com/earth-engine/datasets/catalog/MODIS_006_MCD12Q1#bands)

It turns out that this software is somewhat unsettled.

* steps that depend on rgdal, raster, sp, etc. will be depricated soon
* Version 1.1.3 breaks the mt_to_raster function
  * waiting on feedback from author to solve
  * solved in draft, testing

Not sure how to do legends.

```{r modistools}
# NDVI
products <- mt_products()
View(products)
products[grep("NDVI", products$description),]
bands <- mt_bands(product = "MOD13Q1")
View(bands)
dates <- mt_dates(product = "MOD13Q1", lat = 32, lon = -114)
View(dates)
dates_june2august_ndvi = dates %>%
  filter(month(calendar_date) %in% 6:8)

# Land cover
products[grep("LC", products$description),]
bands_LC <- mt_bands(product = "MCD12Q1")
View(bands_LC)
dates_LC <- mt_dates(product = "MCD12Q1", lat = 32, lon = -114)
View(dates_LC)
dates_june2august_ndvi = dates_LC %>%
  filter(month(calendar_date) %in% 6:8)


# Set some example points
testpoints = data.frame(
  Point_id = c("Old River Channel SFWA", "Willows E of KRP", "Restoration N of KRP HQ"),
  Latitude = c(35.667130, 35.669082, 35.669857),
  Longitude = c(-118.343102, -118.301213, -118.305338))

# test code for one point
tc = mt_subset(lat= 35.667130,
               lon= -118.343102,
                product = "MOD13Q1",
                band = "250m_16_days_NDVI",
                km_lr = 0.7,
                km_ab = 0.7,
                start = "2008-06-09",
                end = "2008-06-09",
                internal = TRUE)

testpoints_sf = st_as_sf(testpoints, coords = c("Longitude", "Latitude"), crs = 4326)
View(testpoints_sf)
plot(st_geometry(testpoints_sf))

# If you need to transform
# tp_6340 = st_transform(testpoints_sf, crs = 6340)

tp_df = sfheaders::sf_to_df(testpoints_sf)
names(tp_df)[3:4] = c("lon", "lat")


# Find date of overflight for use in subsets below
dates08 = mt_dates(product = "MOD13Q1", lat = tp_df$lat[1], lon = tp_df$lon[1])
dates08[lubridate::year(dates08$calendar_date)==2008,]

subsets <- mt_batch_subset(df = tp_df,
                     product = "MOD13Q1",
                     band = "250m_16_days_NDVI",
                     # km_lr = 0.7,
                     # km_ab = 0.7,
                     start = "2008-06-09",
                     end = "2008-06-09",
                     internal = TRUE)


tp_r = mt_to_terra(subsets, reproject = T)

plot(tp_r)
plot(st_geometry(testpoints_sf), add = T)

# Find date of overflight for use in subsets below
dates08_lc = mt_dates(product = "MCD12Q1", lat = tp_df$lat[1], lon = tp_df$lon[1])
dates08_lc[lubridate::year(dates08_lc$calendar_date)==2008,]

subsets_lc <- mt_batch_subset(df = tp_df,
                     product = "MCD12Q1",
                     band = "LC_Type1",
                     # km_lr = 0.7,
                     # km_ab = 0.7,
                     start = "2008-01-01",
                     end = "2008-01-01",
                     internal = TRUE)

tp_r_lc = mt_to_terra(subsets_lc, reproject = T)

subsets_lc <- mt_batch_subset(df = tp_df,
                     product = "MCD12Q1",
                     band = "LC_Prop1",
                     km_lr = 0.7,
                     km_ab = 0.7,
                     start = "2008-01-01",
                     end = "2008-01-01",
                     internal = TRUE)

tp_r_lc = mt_to_terra(subsets_lc, reproject = T)

plot(tp_r_lc)
text(tp_r_lc)
# tp_r_lc@legend
plot(st_geometry(testpoints_sf), add = T)

```


## Using MODIStsp

This now works, but continues to depend on many of the soon to be unsupported packages.

```{r modistsp}
MODIStsp()

```

## Using terr and luna

Based on this example:
https://rspatial.org/modis/2-download.html

```{r terraluna}
prod <- getProducts()

prod_ndvi = prod %>%
  filter(grepl("(?i)NDVI", x = short_name))

product = "US_MODIS_NDVI_1299"
product = "dsp10_ndvi_600"
product = "GOSAT_Level3_of_TANSO-CAI_NDVI"
product = 
product = 
product = 
productInfo(product)
```


## PRISM weather data

https://github.com/ropensci/prism

We may want to group these by the 12 months before the

### Bioclim summary

https://www.worldclim.org/data/bioclim.html

Summarizing weather data by month and quarter according to Bioclim method.

Layer descriptions:

BIO1 = Annual Mean Temperature
BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
BIO3 = Isothermality (BIO2/BIO7) (×100)
BIO4 = Temperature Seasonality (standard deviation ×100)
BIO5 = Max Temperature of Warmest Month
BIO6 = Min Temperature of Coldest Month
BIO7 = Temperature Annual Range (BIO5-BIO6)
BIO8 = Mean Temperature of Wettest Quarter
BIO9 = Mean Temperature of Driest Quarter
BIO10 = Mean Temperature of Warmest Quarter
BIO11 = Mean Temperature of Coldest Quarter
BIO12 = Annual Precipitation
BIO13 = Precipitation of Wettest Month
BIO14 = Precipitation of Driest Month
BIO15 = Precipitation Seasonality (Coefficient of Variation)
BIO16 = Precipitation of Wettest Quarter
BIO17 = Precipitation of Driest Quarter
BIO18 = Precipitation of Warmest Quarter
BIO19 = Precipitation of Coldest Quarter

```{r prism}
prism_get_dl_dir()
# prism_set_dl_dir("Prism")
prism_check_dl_dir()

# This may take a long time on a slow connection (not bad with new Starlink)
get_prism_monthlys(type = "ppt", year = 2021:2022, mon = 1:12, keepZip = FALSE)
get_prism_monthlys(type = "tmax", year = 2021:2022, mon = 1:12, keepZip = FALSE)
get_prism_monthlys(type = "tmin", year = 2021:2022, mon = 1:12, keepZip = FALSE)

prism_archive_ls()

ppt_21 = prism_archive_subset("ppt", "monthly", years = 2021)
# ppt_22 = prism_archive_subset("ppt", "monthly", years = 2022)
ppt_21_7_12 = prism_archive_subset("ppt", "monthly", mon = 7:12, years = 2021)
ppt_22_1_6 = prism_archive_subset("ppt", "monthly", mon = 1:6, years = 2022)

tmax_21 = prism_archive_subset("tmax", "monthly", years = 2021)
# tmax_22 = prism_archive_subset("tmax", "monthly", years = 2022)
tmax_21_7_12 = prism_archive_subset("tmax", "monthly", mon = 7:12, years = 2021)
tmax_22_1_6 = prism_archive_subset("tmax", "monthly", mon = 1:6, years = 2022)

tmin_21 = prism_archive_subset("tmin", "monthly", years = 2021)
# tmin_22 = prism_archive_subset("tmin", "monthly", years = 2022)
tmin_21_7_12 = prism_archive_subset("tmin", "monthly", mon = 7:12, years = 2021)
tmin_22_1_6 = prism_archive_subset("tmin", "monthly", mon = 1:6, years = 2022)

ppt_21 = pd_to_file(ppt_21)
# ppt_22 = pd_to_file(ppt_22)
ppt_21_7_12= pd_to_file(ppt_21_7_12)
ppt_22_1_6 = pd_to_file(ppt_22_1_6)

tmax_21 = pd_to_file(tmax_21)
# tmax_22 = pd_to_file(tmax_22)
tmax_21_7_12 = pd_to_file(tmax_21_7_12)
tmax_22_1_6 = pd_to_file(tmax_22_1_6)

tmin_21 = pd_to_file(tmin_21)
# tmin_22 = pd_to_file(tmin_22)
tmin_21_7_12 = pd_to_file(tmin_21_7_12)
tmin_22_1_6 = pd_to_file(tmin_22_1_6)

# dismo::biovars does not work on spatrasters
#   But maybe someday...
# ppt_21 = rast(ppt_21)

# Get period of interest all of 2021 and months 1-6 of 2023
ppt = raster::stack(c(ppt_21, ppt_22_1_6))
tmax = raster::stack(c(tmax_21, tmax_22_1_6))
tmin = raster::stack(c(tmin_21, tmin_22_1_6))

# ppt_21 = raster::stack(ppt_21)
# ppt_22 = raster::stack(ppt_22)
# tmax_21 = raster::stack(tmax_21)
# tmax_22 = raster::stack(tmax_22)
# tmin_21 = raster::stack(tmin_21)
# tmin_22 = raster::stack(tmin_22)

# needs to be 12 months for biovars
ppt12 = raster::stack(c(ppt_21_7_12, ppt_22_1_6))
tmax12 = raster::stack(c(tmax_21_7_12, tmax_22_1_6))
tmin12 = raster::stack(c(tmin_21_7_12, tmin_22_1_6))
biovars = biovars(prec = ppt12, tmin = tmin12, tmax = tmax12)
crs(biovars)

biovarsprism = "C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/YBCU - Rangewide Surveys CSWG/GIS/PrismWeatherSummaries2021_2022"

# Multiband raster
raster::writeRaster(biovars, filename = file.path(biovarsprism, "biovarsPrismJune21_22.bil"), format = "EHdr",overwrite = T)
# Bands seperated
raster::writeRaster(biovars, filename = file.path(biovarsprism, "biovarsPrismJune21_22.bil"), format = "EHdr",overwrite = T, bylayer = T)

```

## summarize data by sub-watersheds

'hybas_na_lev01-12_v1c' has watersheds at 12 levels, so this approach would work at other levels.

```{r subwatersheds}
huc8 = readOGR("hybas_na_lev01-12_v1c/hybas_na_lev08_v1c.shp")

huc8_2 = huc8

# Mean
# spatRaster method
huc8_vect = vect("hybas_na_lev01-12_v1c/hybas_na_lev08_v1c.shp")
ppt_sr = rast(ppt)
proj4string(ppt)
crs(ppt_sr) = "epsg:4269"
huc8_vect_4269 = project(huc8_vect, ppt_sr)
ppt_mean_huc8_vect = terra::extract(ppt_sr, huc8_vect_4269, fun = mean, bind = T)
writeVector(ppt_mean_huc8_vect, 
            filename = file.path(biovarsprism, "ppt_mean_huc8_vect.shp"), 
            filetype = "ESRI Shapefile", 
            overwrite=TRUE)

# raster method
ppt_mean_huc8 = extract(ppt, huc8, fun = mean)

ppt_mean_huc8.df = as.data.frame(ppt_mean_huc8)
huc8@data = data.frame(huc8@data, ppt_mean_huc8.df)

colnames(huc8@data)[14:31] = c(paste("pptmn21", 1:12, sep = "_"), 
                               paste("pptmn22", 1:6, sep = "_"))
writeOGR(huc8, ".", "ppt_mean_huc8.shp", driver = "ESRI Shapefile")
writeOGR(huc8, file.path(biovarsprism, "ppt_mean_huc8.shp"), ppt_mean_huc8, driver = "ESRI Shapefile")

# Sum
# spatRaster method
# huc8_vect = vect("hybas_na_lev01-12_v1c/hybas_na_lev08_v1c.shp")
# ppt_sr = rast(ppt)
ppt_sum_huc8_vect = terra::extract(ppt_sr, huc8_vect_4269, fun = sum, bind = T)
writeVector(ppt_sum_huc8_vect, 
            filename = file.path(biovarsprism, "ppt_sum_huc8_vect.shp"), 
            filetype = "ESRI Shapefile", 
            overwrite=TRUE)

# raster method
# Not used
# ppt_sum_huc8 = extract(ppt, huc8, fun = sum)
# 
# ppt_sum_huc8.df = as.data.frame(ppt_sum_huc8)
# 
# huc8_2@data = data.frame(huc8_2@data, ppt_sum_huc8.df)
# 
# colnames(huc8_2@data)[14:31] = c(paste("pptsum21", 1:12, sep = "_"), 
#                                paste("pptsum22", 1:6, sep = "_"))
# writeOGR(huc8_2, ".", fname = file.path(biovarsprism, "ppt_sum_huc8.shp"), driver = "ESRI Shapefile")

```



## Some better plots

### Leaflet

From here:
https://rspatialdata.github.io/vegetation.html

```{r leaflet}
label_ndvi = "NDVI"
tp_r_ndvi = raster(mt_to_terra(subsets, reproject = T))

label_lc = "Land Cover Class"
tp_r_lc = raster(mt_to_terra(subsets_lc, reproject = T))

# Defining color palette
pal_ndvi <- colorNumeric(c("#440154FF", "#238A8DFF", "#FDE725FF"), values(tp_r_ndvi), na.color = "transparent")
pal_lc <- colorNumeric(c("#440154FF", "#238A8DFF", "#FDE725FF"), values(tp_r_lc), na.color = "transparent")

# Visualising using leaflet
leaflet() %>%
  addTiles() %>%
  addRasterImage(tp_r_ndvi, colors = pal_ndvi, opacity = 0.75, group = label_ndvi) %>%
  # addRasterImage(tp_r_lc, colors = pal_lc, opacity = 0.75, group = label_lc) %>%
  addLegend(
    pal = pal_ndvi, values = values(tp_r_ndvi),
    title = label_ndvi
  )  %>% 
  # addLegend(
  #   pal = pal_lc, values = values(tp_r_lc),
  #   title = label_lc
  # )  %>% 
  addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
  addProviderTiles("Esri.WorldImagery", group = "ESRI World Imagery") %>%
  addProviderTiles("Esri.WorldTopoMap", group = "ESRI World Topo") %>%
  addLayersControl(
    baseGroups = c("OpenStreetMap", "ESRI World Imagery", "ESRI World Topo"),
    # overlayGroups = c(label_ndvi, label_lc),
    overlayGroups = c(label_ndvi),
    options = layersControlOptions(collapsed = FALSE,),
    position = "bottomleft"
  )

   
```


