---
title: "PrismDataGrab"
author: "Patrick D. lorch"
date: "2024-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## PRISM weather data

https://github.com/ropensci/prism

We may want to group these by the 12 months before the

### Bioclim summary

https://www.worldclim.org/data/bioclim.html

Summarizing weather data by month and quarter according to Bioclim method.

Layer descriptions:

BIO1 = Annual Mean Temperature
BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
BIO3 = Isothermality (BIO2/BIO7) (×100)
BIO4 = Temperature Seasonality (standard deviation ×100)
BIO5 = Max Temperature of Warmest Month
BIO6 = Min Temperature of Coldest Month
BIO7 = Temperature Annual Range (BIO5-BIO6)
BIO8 = Mean Temperature of Wettest Quarter
BIO9 = Mean Temperature of Driest Quarter
BIO10 = Mean Temperature of Warmest Quarter
BIO11 = Mean Temperature of Coldest Quarter
BIO12 = Annual Precipitation
BIO13 = Precipitation of Wettest Month
BIO14 = Precipitation of Driest Month
BIO15 = Precipitation Seasonality (Coefficient of Variation)
BIO16 = Precipitation of Wettest Quarter
BIO17 = Precipitation of Driest Quarter
BIO18 = Precipitation of Warmest Quarter
BIO19 = Precipitation of Coldest Quarter

### Get new prism data



```{r prism}
library(prism)
# library(dismo) #Does not work. Still dependent on raster and sp
library(terra)
library(luna)

prism_get_dl_dir()
prism_set_dl_dir("Prism")
prism_check_dl_dir()

# This may take a long time on a slow connection (not bad with new Starlink)
get_prism_monthlys(type = "ppt", year = 2023:2024, mon = 1:12, keepZip = FALSE)
get_prism_monthlys(type = "tmax", year = 2023:2024, mon = 1:12, keepZip = FALSE)
get_prism_monthlys(type = "tmin", year = 2023:2024, mon = 1:12, keepZip = FALSE)

prism_archive_ls()

ppt_21 = prism_archive_subset("ppt", "monthly", years = 2021)
ppt_22 = prism_archive_subset("ppt", "monthly", years = 2022)
ppt_23 = prism_archive_subset("ppt", "monthly", years = 2023)
ppt_24 = prism_archive_subset("ppt", "monthly", years = 2024)
# These methods get 12 months before June to be used as covariates for 
#   bird breeding or NDVI analyses
# ppt_21_7_12 = prism_archive_subset("ppt", "monthly", mon = 7:12, years = 2021)
# ppt_22_1_6 = prism_archive_subset("ppt", "monthly", mon = 1:6, years = 2022)

tmax_21 = prism_archive_subset("tmax", "monthly", years = 2021)
tmax_22 = prism_archive_subset("tmax", "monthly", years = 2022)
tmax_23 = prism_archive_subset("tmax", "monthly", years = 2023)
tmax_24 = prism_archive_subset("tmax", "monthly", years = 2024)
# tmax_21_7_12 = prism_archive_subset("tmax", "monthly", mon = 7:12, years = 2021)
# tmax_22_1_6 = prism_archive_subset("tmax", "monthly", mon = 1:6, years = 2022)

tmin_21 = prism_archive_subset("tmin", "monthly", years = 2021)
tmin_22 = prism_archive_subset("tmin", "monthly", years = 2022)
tmin_23 = prism_archive_subset("tmin", "monthly", years = 2023)
tmin_24 = prism_archive_subset("tmin", "monthly", years = 2024)
# tmin_21_7_12 = prism_archive_subset("tmin", "monthly", mon = 7:12, years = 2021)
# tmin_22_1_6 = prism_archive_subset("tmin", "monthly", mon = 1:6, years = 2022)

ppt_21 = pd_to_file(ppt_21)
ppt_22 = pd_to_file(ppt_22)
ppt_23 = pd_to_file(ppt_23)
ppt_24 = pd_to_file(ppt_24)
# ppt_21_7_12= pd_to_file(ppt_21_7_12)
# ppt_22_1_6 = pd_to_file(ppt_22_1_6)

tmax_21 = pd_to_file(tmax_21)
tmax_22 = pd_to_file(tmax_22)
tmax_23 = pd_to_file(tmax_23)
tmax_24 = pd_to_file(tmax_24)
# tmax_21_7_12 = pd_to_file(tmax_21_7_12)
# tmax_22_1_6 = pd_to_file(tmax_22_1_6)

tmin_21 = pd_to_file(tmin_21)
tmin_22 = pd_to_file(tmin_22)
tmin_23 = pd_to_file(tmin_23)
tmin_24 = pd_to_file(tmin_24)
# tmin_21_7_12 = pd_to_file(tmin_21_7_12)
# tmin_22_1_6 = pd_to_file(tmin_22_1_6)

# dismo::biovars does not work on spatrasters
#   But maybe someday...
# ppt_21 = rast(ppt_21)

# Get period of interest all of 2021 and months 1-6 of 2022
# This is too many months for bioclim summaries
# ppt = raster::stack(c(ppt_21, ppt_22_1_6))
# tmax = raster::stack(c(tmax_21, tmax_22_1_6))
# tmin = raster::stack(c(tmin_21, tmin_22_1_6))

ppt_21 = raster::stack(ppt_21)
ppt_22 = raster::stack(ppt_22)
ppt_23 = raster::stack(ppt_23)
ppt_24 = raster::stack(ppt_24)
tmax_21 = raster::stack(tmax_21)
tmax_22 = raster::stack(tmax_22)
tmax_23 = raster::stack(tmax_23)
tmax_24 = raster::stack(tmax_24)
tmin_21 = raster::stack(tmin_21)
tmin_22 = raster::stack(tmin_22)
tmin_23 = raster::stack(tmin_23)
tmin_24 = raster::stack(tmin_24)

# needs to be 12 months for biovars
# ppt12 = raster::stack(c(ppt_21_7_12, ppt_22_1_6))
# tmax12 = raster::stack(c(tmax_21_7_12, tmax_22_1_6))
# tmin12 = raster::stack(c(tmin_21_7_12, tmin_22_1_6))
biovars_21 = biovars(prec = ppt_21, tmin = tmin_21, tmax = tmax_21)
crs(biovars_21)
biovars_22 = biovars(prec = ppt_22, tmin = tmin_22, tmax = tmax_22)
crs(biovars_22)
biovars_23 = biovars(prec = ppt_23, tmin = tmin_23, tmax = tmax_23)
crs(biovars_23)
# Dec not complete at time of running so not doing this one
# biovars_24 = biovars(prec = ppt_24, tmin = tmin_24, tmax = tmax_24)
# crs(biovars_24)

dir.create("PrismRasters")
biovarsprism = "PrismRasters"
# biovarsprism = "C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/YBCU - Rangewide Surveys CSWG/GIS/PrismWeatherSummaries2021_2022"

# Multiband raster
# raster::writeRaster(biovars_21, filename = file.path(biovarsprism, "biovarsPrism21.bil"), format = "EHdr",overwrite = T)
# Bands seperated
raster::writeRaster(biovars_21, filename = file.path(biovarsprism, "biovarsPrism21.bil"), format = "EHdr",overwrite = T, bylayer = T)

# raster::writeRaster(biovars_22, filename = file.path(biovarsprism, "biovarsPrism22.bil"), format = "EHdr",overwrite = T)
# Bands seperated
raster::writeRaster(biovars_22, filename = file.path(biovarsprism, "biovarsPrism22.bil"), format = "EHdr",overwrite = T, bylayer = T)

# raster::writeRaster(biovars_23, filename = file.path(biovarsprism, "biovarsPrism23.bil"), format = "EHdr",overwrite = T)
# Bands seperated
raster::writeRaster(biovars_23, filename = file.path(biovarsprism, "biovarsPrism23.bil"), format = "EHdr",overwrite = T, bylayer = T)

# Dec not complete at time of running so not doing this one
# raster::writeRaster(biovars_24, filename = file.path(biovarsprism, "biovarsPrism24.bil"), format = "EHdr",overwrite = T)
# # Bands seperated
# raster::writeRaster(biovars_24, filename = file.path(biovarsprism, "biovarsPrism24.bil"), format = "EHdr",overwrite = T, bylayer = T)

```

## summarize data by sub-watersheds

'hybas_na_lev01-12_v1c' has watersheds at 12 levels, so this approach would work at other levels.

```{r subwatersheds}
huc8 = readOGR("hybas_na_lev01-12_v1c/hybas_na_lev08_v1c.shp")

# Mean
# spatRaster method
huc8_vect = vect("hybas_na_lev01-12_v1c/hybas_na_lev08_v1c.shp")
# plot(huc8_vect)
ppt_21_sr = rast(ppt_21)
proj4string(ppt_21)
crs(ppt_21_sr) = "epsg:4269"
huc8_vect_4269 = project(huc8_vect, ppt_21_sr)
ppt_21_mean_huc8_vect = terra::extract(ppt_21_sr, huc8_vect_4269, fun = mean, bind = T)
# writeVector(ppt_21_mean_huc8_vect, 
#             filename = file.path(biovarsprism, "ppt_21_mean_huc8_vect.shp"), 
#             filetype = "ESRI Shapefile", 
#             overwrite=TRUE)

ppt_21_mean_huc8_vect$mean_yr = mean(ppt_21_mean_huc8_vect[,14:25])

# raster method
ppt_mean_huc8 = extract(ppt, huc8, fun = mean)

ppt_mean_huc8.df = as.data.frame(ppt_mean_huc8)
huc8@data = data.frame(huc8@data, ppt_mean_huc8.df)

colnames(huc8@data)[14:31] = c(paste("pptmn21", 1:12, sep = "_"), 
                               paste("pptmn22", 1:6, sep = "_"))
writeOGR(huc8, ".", "ppt_mean_huc8.shp", driver = "ESRI Shapefile")
writeOGR(huc8, file.path(biovarsprism, "ppt_mean_huc8.shp"), ppt_mean_huc8, driver = "ESRI Shapefile")

# Sum
# spatRaster method
# huc8_vect = vect("hybas_na_lev01-12_v1c/hybas_na_lev08_v1c.shp")
# ppt_sr = rast(ppt)
ppt_21_sum_huc8_vect = terra::extract(ppt_21_sr, huc8_vect_4269, fun = sum, bind = T)
# writeVector(ppt_21_sum_huc8_vect, 
#             filename = file.path(biovarsprism, "ppt_sum_huc8_vect.shp"), 
#             filetype = "ESRI Shapefile", 
#             overwrite=TRUE)

# raster method
# Not used
# ppt_sum_huc8 = extract(ppt, huc8, fun = sum)
# 
# ppt_sum_huc8.df = as.data.frame(ppt_sum_huc8)
# 
# huc8_2@data = data.frame(huc8_2@data, ppt_sum_huc8.df)
# 
# colnames(huc8_2@data)[14:31] = c(paste("pptsum21", 1:12, sep = "_"), 
#                                paste("pptsum22", 1:6, sep = "_"))
# writeOGR(huc8_2, ".", fname = file.path(biovarsprism, "ppt_sum_huc8.shp"), driver = "ESRI Shapefile")

```

## Plots to check extents and look at rasters

```{r simpleplots}
library(ggplot2)

caltowhee_ext = ext(-118.000, -117.120, 35.5, 36.25)
caltowhee_points = data.frame(longitude = c(-117.53839, -117.39282),
                              latitude = c(35.94762, 35.96628))

# BIO13 =  Precipitation of Wettest Month
plot(terra::crop(rast(biovars_21$bio13), caltowhee_ext))
lines(ppt_21_mean_huc8_vect)
points(caltowhee_points, pch = 19)
# BIO16 = Precipitation of Wettest Quarter
plot(terra::crop(rast(biovars_21$bio16), caltowhee_ext))
lines(ppt_21_mean_huc8_vect)
# BIO19 = Precipitation of Coldest Quarter
plot(terra::crop(rast(biovars_21$bio19), caltowhee_ext))
lines(ppt_21_mean_huc8_vect)

# BIO13 =  Precipitation of Wettest Month
ppt_wet_mnth_21 = biovars_21$bio13 %>%
  raster::extract(caltowhee_points)
ppt_wet_mnth_22 = biovars_22$bio13 %>%
  raster::extract(caltowhee_points)
ppt_wet_mnth_23 = biovars_23$bio13 %>%
  raster::extract(caltowhee_points)
ppt_wet_mnth_21_23 = data.frame(rbind(ppt_wet_mnth_21, ppt_wet_mnth_22, ppt_wet_mnth_23))
names(ppt_wet_mnth_21_23) = c("TowheeNAWS1", "TowheeBLM1")
ppt_wet_mnth_21_23$Year = 21:23
p13 = ggplot(ppt_wet_mnth_21_23, aes(Year)) + 
  geom_line(aes(y = TowheeNAWS1), col = "grey") +
  geom_line(aes(y = TowheeBLM1), col = "red")
p13

# BIO16 = Precipitation of Wettest Quarter
ppt_wet_qrtr_21 = biovars_21$bio16 %>%
  raster::extract(caltowhee_points)
ppt_wet_qrtr_22 = biovars_22$bio16 %>%
  raster::extract(caltowhee_points)
ppt_wet_qrtr_23 = biovars_23$bio16 %>%
  raster::extract(caltowhee_points)
ppt_wet_qrtr_21_23 = data.frame(rbind(ppt_wet_qrtr_21, ppt_wet_qrtr_22, ppt_wet_qrtr_23))
names(ppt_wet_qrtr_21_23) = c("TowheeNAWS1", "TowheeBLM1")
ppt_wet_qrtr_21_23$Year = 21:23
p16 = ggplot(ppt_wet_qrtr_21_23, aes(Year)) + 
  geom_line(aes(y = TowheeNAWS1), col = "grey") +
  geom_line(aes(y = TowheeBLM1), col = "red")
p16

# BIO19 = Precipitation of Coldest Quarter
ppt_cold_qrtr_21 = biovars_21$bio19 %>%
  raster::extract(caltowhee_points)
ppt_cold_qrtr_22 = biovars_22$bio19 %>%
  raster::extract(caltowhee_points)
ppt_cold_qrtr_23 = biovars_23$bio19 %>%
  raster::extract(caltowhee_points)
ppt_cold_qrtr_21_23 = data.frame(rbind(ppt_cold_qrtr_21, ppt_cold_qrtr_22, ppt_cold_qrtr_23))
names(ppt_cold_qrtr_21_23) = c("TowheeNAWS1", "TowheeBLM1")
ppt_cold_qrtr_21_23$Year = 21:23
p19 = ggplot(ppt_cold_qrtr_21_23, aes(Year)) + 
  geom_line(aes(y = TowheeNAWS1), col = "grey") +
  geom_line(aes(y = TowheeBLM1), col = "red")
p19

```


## Some better plots

These plots can be exported to an html.  They allow the end user to move around, zoom, and change background images.

### Leaflet

From here:
https://rspatialdata.github.io/vegetation.html

```{r leaflet}
library(dplyr)
library(leaflet)

pal_biovars_21_bio13 = colorNumeric(c("#440154FF", "#238A8DFF", "#FDE725FF"), values(biovars_21$bio13), na.color = "transparent")

# Visualising using leaflet
leaflet() %>%
  addTiles() %>%
  addRasterImage(biovars_21$bio13, 
                 colors = pal_biovars_21_bio13,
                 opacity = 0.75, 
                 group = "Precip., Wettest Mnth.") %>%
  addPolygons(data = huc8, group = "HUC8 watersheds", fillOpacity = 0) %>%
  addCircleMarkers(caltowhee_points$longitude, 
                   caltowhee_points$latitude,
                   color = "red") %>%
  setView(lng = -117.5, lat = 35.9, zoom = 9) %>%
  addLegend(
    pal = pal_biovars_21_bio13, 
    values = values(biovars_21$bio13),
    title = "Precip. (mm)"
  )  %>% 
  addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
  addProviderTiles("Esri.WorldImagery", group = "ESRI World Imagery") %>%
  addProviderTiles("Esri.WorldTopoMap", group = "ESRI World Topo") %>%
  addScaleBar(position = "bottomright") %>%
  addLayersControl(
    baseGroups = c("OpenStreetMap", "ESRI World Imagery", "ESRI World Topo"),
    # overlayGroups = c(label_ndvi, label_lc),
    overlayGroups = c("Precip., Wettest Mnth.",
                      "HUC8 watersheds"),
    options = layersControlOptions(collapsed = TRUE,),
    position = "bottomleft"
  ) 

```